---
title: "ASReview report"
author: "Samuel Langton"
date: "3 January 2022"
output:
  # word_document:
  #   toc: yes
  #   toc_depth: '3'
  html_document:
    toc: yes
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F)
# Load packages.
library(readr)
library(dplyr)
library(stringr)
library(purrr)
library(irr)
library(kableExtra)
```

```{r start}
# Load data.
first_df <- read_csv("data/asreview_result_spatial-and-temporal-patterning-of-emergency-reactive-police-demand.csv")
secon_df <- read_csv("data/asreview_result_spatial-and-temporal-patterning-of-emergency-reactive-police-demand-2nd-coder.csv")

# Total reviewed.
first_total <- table(first_df$included)[[2]] + table(first_df$included)[[1]]
secon_total <- table(secon_df$included)[[2]] + table(secon_df$included)[[1]]

# Relevant flags.
first_rele <- table(first_df$included)[[2]]
first_irre <- table(first_df$included)[[1]]

# Irrelevant flags.
secon_rele <- table(secon_df$included)[[2]]
secon_irre <- table(secon_df$included)[[1]]

# Not reviewed
first_unrev <- sum(is.na(first_df$included))
secon_unrev <- sum(is.na(secon_df$included))

# % reviewed.
first_prev <- round(100*first_total/nrow(first_df), 2)
secon_prev <- round(100*secon_total/nrow(secon_df), 2)

# Total irrelevant.
first_trele <- first_irre + first_unrev
secon_trele <- secon_irre + secon_unrev

# % total relevant.
first_ptrel <- round(100*first_rele/nrow(first_df), 2)
secon_ptrel <- round(100*secon_rele/nrow(secon_df), 2)

# % flagged relevant.
first_pfrel <- round(100*first_rele/first_total, 2)
secon_pfrel <- round(100*secon_rele/secon_total, 2)

# Recode 'not rated' as irrelevant in each data frame.
first_rec_df <- first_df %>%
  mutate(included = if_else(is.na(included), true = 0, false = included))

secon_rec_df <- secon_df %>%
  mutate(included = if_else(is.na(included), true = 0, false = included))


# Disagreements 1: first coder says irrelevant, second coder says relevant.

first_irrelevant_vec <- first_rec_df %>% 
  filter(included == 0) %>% 
  pluck("record_id")

# Here, note that ASReview updated their default column names late-2021. Choose the relevant select() option for your data.
disagree1_df <- secon_rec_df %>% 
  filter(included == 1, record_id %in% first_irrelevant_vec) %>% 
  # select(record_id, authors, year, title, abstract) # For newer projects.
  select(record_id, first_authors, publication_year, primary_title, notes_abstract) # For older projects.

first_dis <- nrow(disagree1_df) # 6

# Disagreement 2: second coder says irrelevant, first coder says relevant.

secon_irrelevant_vec <- secon_rec_df %>% 
  filter(included == 0) %>% 
  pluck("record_id")

# Here, note that ASReview updated their default column names late-2021. Choose the relevant select() option for your data.
disagree2_df <- first_rec_df %>% 
  filter(included == 1, record_id %in% secon_irrelevant_vec) %>% 
  # select(record_id, authors, year, title, abstract) # For newer projects.
  select(record_id, first_authors, publication_year, primary_title, notes_abstract) # For older projects.

secon_dis <- nrow(disagree2_df) # 39

# Sort and create vectors for Kappa.
first_vec <- first_rec_df %>% 
  arrange(record_id) %>% 
  pluck("included")

secon_vec <- secon_rec_df %>% 
  arrange(record_id) %>% 
  pluck("included")

# Create array for Kappa.
full_array <- cbind(first_vec, secon_vec)

# Further disgareement statistics.

# Identify those which each reviewer did not even look at.
first_dist_vec <- first_rec_df %>% 
  filter(included == 3) %>% 
  pluck("record_id")

secon_dist_vec <- secon_rec_df %>% 
  filter(included == 3) %>% 
  pluck("record_id")

# Identify those relevant by one rather, and unreviewed by the other.
first_dist_df <- first_rec_df %>% 
  filter(included == 1) %>% 
  mutate(missed = if_else(record_id %in% secon_dist_vec,
                          "missed",
                          "not missed"))

secon_dist_df <- secon_rec_df %>% 
  filter(included == 1) %>% 
  mutate(missed = if_else(record_id %in% secon_dist_vec,
                          "missed",
                          "not missed"))

# How many studies did one rather say *relevcant* and the other reviewer
# not even review?
missed1 <- nrow(first_dist_df) - table(first_dist_df$missed)[[1]]
missed2 <- nrow(secon_dist_df) - table(secon_dist_df$missed)[[1]]
```

# Summary

|                          | rater 1            | rater 2            |
| ------------------------ |:------------------:| ------------------:|
| n uploaded               | `r nrow(first_df)` | `r nrow(secon_df)` |
| n reviewed               | `r first_total`    | `r secon_total`    |
| % reviewed               | `r first_prev`     | `r secon_prev`     |
| n flagged relevant       | `r first_rele`     | `r secon_rele`     |
| n flagged irrelevant     | `r first_irre`     | `r secon_irre`     |
| % flagged relevant       | `r first_pfrel`    | `r secon_pfrel`    |
| n unreviewed             | `r first_unrev`    | `r secon_unrev`    |
| n total irrelevant       | `r first_trele`    | `r secon_trele`    |
| % total relevant         | `r first_ptrel`    | `r secon_ptrel`    |
| n relevant v. irrelevant | `r first_dis`      | `r secon_dis`      | 
| n relevant v. unreviewed | `r missed1`        | `r missed2`        | 

```{r irrdatasave}
irr_df <- tibble(
  ` ` = c("n uploaded", 
                "n reviewed",
                "% reviewed",
                "n flagged relevant",
                "n flagged irrelevant",
                "% flagged relevant",
                "n unreviewed",
                "n total irrelevant",
                "% total relevant",
                "n relevant v. irrelevant",
                "n relevant v. unreviewed"),
  `rater 1` = c(nrow(first_df),
                first_total,
                first_prev,
                first_rele,
                first_irre,
                first_pfrel,
                first_unrev,
                first_trele,
                first_ptrel,
                first_dis,
                missed1),
  `rater 2` = c(nrow(secon_df),
                secon_total,
                secon_prev,
                secon_rele,
                secon_irre,
                secon_pfrel,
                secon_unrev,
                secon_trele,
                secon_ptrel,
                secon_dis,
                missed2)
  )

write_csv(x = irr_df, file = "data/irr_table.csv")

```

# Kappa

The level of agreement between rater 1 and rater 2, assuming that unreviewed articles are irrelevant, is the following:

```{r}
agree(full_array)
```

The Kappa statistic is the following:

```{r}
kappa2(full_array)
```

# Disagreements

## Rater 1

Studies for which  rater 1 flagged relevant and rater 2 flagged irrelevant.

```{r}
# Here, note that ASReview updated their default column names late-2021. 
# If your project is newer (late-2021 onward) then comment out the mutate() line.

disagree2_df %>% 
  mutate(publication_year = str_remove_all(publication_year, "///")) %>% # Only needed for older projects.
  kable(col.names = c("id", "author(s)", "year", "title", "abstract")) %>%
  # scroll_box(width = "120%", height = "300px") %>% 
  kable_styling()
```

<br>

<br>

## Rater 2

Studies for which  rater 2 flagged relevant and rater 1 flagged irrelevant.

```{r}
# Here, note that ASReview updated their default column names late-2021. 
# If your project is newer (late-2021 onward) then comment out the mutate() line.

disagree1_df %>% 
  mutate(publication_year = str_remove_all(publication_year, "///")) %>% # Only needed for older projects.
  kable(col.names = c("id", "author(s)", "year", "title", "abstract")) %>% 
  # scroll_box(width = "120%", height = "300px") %>% 
  kable_styling()
```

